Oregon 2013 Medication for Beta-2 agonists

### 2017-07-18 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

---
title: "Oregon 2013 Medication for Asthma Inhalers"
author: "Jingyang Liu"
date: "May 17, 2017"
output: html_document
---

@@

# two-stage distributed lag model libraries
library(dlnm)
library(mvmeta)
library(splines)

@@

We had previously shown GWR to be a good estimate of wildfire smoke concentrations, and showed associations with respiratory outcomes (including asthma) in Washington state. We want to replicate results for Oregon and also look at beta2 agonist pharmacy fills.

For this research I have the Oregon All Payer All Claims (APAC) data set in 2013. I want to research on the effects of wildfire smoke on beta-2 agonist pharmacy fill, a medication primarily used as a rescue medication for persons with asthma or COPD. NDC (National Drug Code) is the variable in the data set that the data set offered. The standard NDC I use the link below, which is a ndc list in 2013, which fits my original data set. 
(http://ww2.ncqa.org/hedis-quality-measurement/hedis-measures/hedis-2013/hedis-2013-final-ndc-lists)

The data cleaning and creation of the casecross-over dataframes are huge computation tasks, so they are done with server, the following work is done with local R.

@@

plot_smk_ts <- ggplot(or_ndc_county_ts, aes(y =geo_smk_pm10, x = date)) +
           geom_jitter() +
           ylab("GWR smoke concentration (10ug/m^3)") +
           xlab("Year 2013")

print(plot_smk_ts)

@@

```{r}
meta_ts <- read_csv("../data/meta_timeseries.csv") %>% 
  # creating 10 unit increase var of geo smoke and day of week
  mutate(geo_smk10 = geo_smk_pm/10,
         day = weekdays(date)) %>% 
  # create smoke and lag variables
  mutate(smoke0 = ifelse(geo_smk_pm > 0, 1, 0),
         smoke5 = ifelse(geo_smk_pm > 5, 1, 0),
         smoke10 = ifelse(geo_smk_pm > 10, 1, 0),
         smoke15 = ifelse(geo_smk_pm > 15, 1, 0), 
         # lag variable
         geo_smk_lag1 = lag(geo_smk10, 1, order_by = county),
         geo_smk_lag2 = lag(geo_smk10, 2, order_by = county),
         geo_smk_lag3 = lag(geo_smk10, 3, order_by = county),
         geo_smk_lag4 = lag(geo_smk10, 4, order_by = county),
         geo_smk_lag5 = lag(geo_smk10, 5, order_by = county),
         geo_smk_lag6 = lag(geo_smk10, 6, order_by = county),
         geo_smk_lag7 = lag(geo_smk10, 7, order_by = county)) %>% 
  ungroup()

# smoke wave days (at least 2 consecutive days in the 98th percentile of PM)
quantile(meta_ts$geo_wt_pm, probs = c(0.95,0.98,0.99, 0.999))
# following Coco's smoke wave day, the 98th percentile for PM2.5 concentrations
# would be >26.9

smoke_wave <- meta_ts %>% 
  select(county, date, geo_wt_pm) %>% 
  mutate(high_pm_day = ifelse(geo_wt_pm > 26.9, 1, 0)) %>% 
  group_by(county) %>% 
  # identify consecutive days
  mutate(smoke_wave = ifelse(lag(high_pm_day, 
    order_by = county)==1 & high_pm_day ==1, 1, 0)) %>% 
  select(-high_pm_day, -geo_wt_pm)

# merge in smokewave day
meta_ts <- meta_ts %>% 
  right_join(smoke_wave, by = c("county", "date"))

# count of state outcome timeseries
state_outcome_rate <- meta_ts %>% 
  group_by(date) %>% 
  # remove columns I don't want to calculate
  select(-fips, -county) %>% 
  # add up outcomes in each county
  summarise_each(funs(sum), population, n_obs:broken_arm_n) %>% 
  mutate_each(funs((./population*100000)), n_obs:broken_arm_n)

# average of environmental measures, joined with outcomes
state_smoke <- meta_ts %>% 
  group_by(state, date) %>% 
  select(-fips, -county) %>% 
  summarise_each(funs(mean, median, min, max), wrf_pm:wrf_temp) %>% 
  full_join(state_outcome_rate, by = c("date"))

# only 150 smoke wave days
xtabs(~smoke_wave,  meta_ts)
# if we compare to our geo smoke > 10, we'd add 362 extra days
xtabs(~smoke_wave + smoke10, meta_ts)
# if we used smoke >15 we'd add 164 extra days 
xtabs(~smoke_wave + smoke15, meta_ts)

# small multiples dataset
or_smk <- or_ts %>% 
  select(county, date, wrf_smk_pm, krig_smk_pm, geo_smk_pm) %>% 
  gather(key = smk_method, value = pm, -date, -county)

@@

# data wrangle ----
#first step is to summarise the counts of smoke days in each county 
or_smk_count <- meta_ts %>% 
  group_by(county) %>% 
  # summing up smoke days
  summarise(smk_wave_n = sum(smoke_wave), smk_day0 = sum(smoke0), 
            smk_day5 = sum(smoke5), smk_day10 = sum(smoke10),
            smk_day15 = sum(smoke15)) %>% 
  # county needs to be lowercase to join to maps dataframe
  mutate(county = tolower(county))

@@

# outcome name
outcome_name <- "beta-2 agonist"

@@

@@

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

@@

# outcome name
outcome <- "Beta-2 agonists"
outcome_name <- "Beta-2 agonists"

@@

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- inhaler_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases

@@

# empty matrix (12 x 10 matrix)
point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

@@

# dataframe for analysis creation
# bind columns back together 
df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
  # remove missing pm values
  filter(!is.na(wrf_smk_pm_zip)) %>% 
  # the following code makes sure that the counterfactual values retained are 
  # symetric in that number of obs before = number of obs after
  mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

# second loop to run a model for each pm estimation method
for(j in 38:40){

@@

  # create else statement that fills matrix with missing so I still have the row
  # in the final dataframe
  } else {point_estimates[row_n, 2] <- method_list[row_n]
          point_estimates[row_n, 3] <- 0
          point_estimates[row_n, 4] <- 0
          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
  } # end methods loop

@@

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" |
    #      pm_method == "WRF-Chem Smoke") %>% 
  # subset columns I want to put in to the table
  select(2, 4:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
         # row group by outcome
         rgroup = "Asthma Inhalers",
         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Est Method", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(3, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
         ) # end table

@@

# outcome name
outcome <- "asthma inhalers"
outcome_name <- "Asthma inhalers"

@@

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- inhaler_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases

# new loop for age categories
for(k in 0:2){
  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

@@

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_zip)) %>% 
    # limit to specific age category
    filter(age_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
    # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
  for(j in 38:40){

@@

    # create else statement that fills matrix with missing so I still have the row
    # in the final dataframe
    } else {point_estimates[row_n, 3] <- method_list[row_n]
            point_estimates[row_n, 4] <- 0
            point_estimates[row_n, 5] <- 0
            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
    } # end methods loop

@@

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
         # row group by outcome
         rgroup = "Asthma inhalers",
         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Est Method", "Age Group", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(4, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
         ) # end table

@@

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_age_fig_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

@@

Association with smoke PM~2.5~ and health outcomes stratified by sex.

@@

These results show that there may be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.

```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

@@

  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

@@

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_zip)) %>% 
    # limit to specific sex category
    filter(sex_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
    # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
  for(j in 38:40){

@@

  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
           # row group by outcome
           rgroup = "Asthma inhalers",
           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
           # column headers
           header = c("Est Method", "Sex Strata", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc", # column alignment,
           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
            ) # end table

@@

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_sex_fig_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

@@

Remove the missing values for GWR smoke. But I only use GWR smoke of zip for the case crossover study.

@@

# fill outcome name in first column
fit_mat[,1] <- "inhalers"
# fill in exposure (geo smoke)
fit_mat[,2] <- "geo_smk10"
# fit type
fit_mat[,3] <- "ns"

@@

@@

# kable
knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline",
  "degree of freedom best fit by AIC for mixed model approach"))
```

@@

```{r DLM for casecrossover, echo = FALSE, message=FALSE, warning=FALSE}
inhaler_geo_lag <- inhaler_ndc_lag_df %>%
  select(159, 160:166) # GWR-method

@@

@@

@@

@@

@@

@@

```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE}
read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
or_ndc_claim <- read_csv(read_path7)

@@

```{r time stratified zip for casecrossover lag -7 to 0 df, echo=FALSE, include=FALSE}
read_path10 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
inhaler_ndc_lag15 <- read_csv(read_path10)
# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis

@@

```{r DLM for casecrossover lag -7 to 0, echo = FALSE, message=FALSE, warning=FALSE}
inhaler_geo_lag <- inhaler_ndc_lag_15_df %>%
  select(177:171, 178) # GWR-method

B <- ns(-7:lag_max, df =5, intercept=TRUE)

@@

# outcome name
outcome <- "Asthma Inhalers"

@@

# extract pm values and divide by 10 and ordered
which(colnames(inhaler_ndc)=="geo_smk_pm_lag1_zip") # 136
which(colnames(inhaler_ndc)=="wrf_temp_lag1_zip") # 150
pm_estimates_df <- inhaler_ndc[, 136:142]/10  # create 10 unit increases
temp_estimates_df <- inhaler_ndc[, 150:156]  # temperature

@@

tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
         # row group by outcome
         rgroup = "Asthma Inhalers",
         n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Lag", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(3, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
         ) # end table

@@

rr_df[i,3:5] <- rr_mp
}
rr_df

@@

@@

### 4. Distributed lag -7 to 0 for time series
```{r DLM for casecrossover lag -7 to 0 temp, echo = FALSE, message=FALSE, warning=FALSE, eval = FALSE}

>>>>

---
title: "Oregon 2013 Medication for Beta-2 agonists"
author: "Jingyang Liu"
date: "May 17, 2017"
output: html_document
---

@@

# distributed lag model libraries
library(dlnm)
library(mvmeta)
library(splines)

@@

We had previously shown GWR to be a good estimate of wildfire smoke concentrations, and showed associations with respiratory outcomes (including asthma) in Washington state. We want to replicate results for Oregon and also look at beta-2 agonist pharmacy fills.

For this research I have the Oregon All Payer All Claims (APAC) data set in 2013. I want to research on the effects of wildfire GWR smoke on beta-2 agonist pharmacy fill, a medication primarily used as a rescue medication for persons with asthma or COPD. NDC (National Drug Code) is the variable in the data set that the data set offered. The standard NDC I use the link below, which is a ndc list in 2013 and fits my original data set. 
(http://ww2.ncqa.org/hedis-quality-measurement/hedis-measures/hedis-2013/hedis-2013-final-ndc-lists)

The data cleaning and creation of the casecross-over dataframes are huge computation tasks, so they are done with server, the following work is done with local R. We analyze the case crossover results with overall, age strarified and sex stratified categories for the pharmacy fills, and then we also analyze with the distributed lag method, and check with time series analysis.

@@

# plot_smk_ts <- ggplot(or_ndc_county_ts, aes(y =geo_smk_pm10, x = date)) +
#           geom_jitter() +
#           ylab("GWR smoke concentration (10ug/m^3)") +
#           xlab("Year 2013")

# print(plot_smk_ts)

@@

```{r smoke days and map, warning=FALSE, message=FALSE, echo=FALSE}
or_ts <- or_ndc_county_ts %>% 
  mutate(smoke0 = ifelse(geo_smk_pm_county > 0, 1, 0),
         smoke5 = ifelse(geo_smk_pm_county > 5, 1, 0),
         smoke10 = ifelse(geo_smk_pm_county > 10, 1, 0),
         smoke15 = ifelse(geo_smk_pm_county > 15, 1, 0), 
         # lag variable
         geo_smk_lag1 = lag(geo_smk_pm10, 1, order_by = county),
         geo_smk_lag2 = lag(geo_smk_pm10, 2, order_by = county),
         geo_smk_lag3 = lag(geo_smk_pm10, 3, order_by = county),
         geo_smk_lag4 = lag(geo_smk_pm10, 4, order_by = county),
         geo_smk_lag5 = lag(geo_smk_pm10, 5, order_by = county),
         geo_smk_lag6 = lag(geo_smk_pm10, 6, order_by = county),
         geo_smk_lag7 = lag(geo_smk_pm10, 7, order_by = county)) %>% 
  ungroup()

# smoke wave days (at least 2 consecutive days in the 98th percentile of PM)
# quantile(or_ts$geo_wt_pm_county, probs = c(0.95,0.98,0.99, 0.999))
# following Coco's smoke wave day, the 98th percentile for PM2.5 concentrations
# would be >16.3

smoke_wave <- or_ts %>% 
  select(county, date, geo_wt_pm_county) %>% 
  mutate(high_pm_day = ifelse(geo_wt_pm_county > 16.33, 1, 0)) %>% 
  group_by(county) %>% 
  # identify consecutive days
  mutate(smoke_wave = ifelse(lag(high_pm_day, 
    order_by = county)==1 & high_pm_day ==1, 1, 0)) %>% 
  select(-high_pm_day, -geo_wt_pm_county)

# merge in smokewave day
or_ts <- or_ts %>% 
  right_join(smoke_wave, by = c("county", "date"))

# count of state outcome timeseries
state_outcome_rate <- or_ts %>% 
  group_by(date) %>% 
  # remove columns I don't want to calculate
  select(-fips, -county)

# average of environmental measures, joined with outcomes
state_smoke <- or_ts %>% 
  group_by(date) %>% 
  select(-fips, -county) %>% 
  summarise_each(funs(mean, median, min, max), wrf_pm_county:wrf_temp_county) # %>% 
  # full_join(state_outcome_rate, by = c("date"))

# only 55 smoke wave days
# xtabs(~smoke_wave,  or_ts)
# if we compare to our geo smoke > 10, we'd add 362 extra days
# xtabs(~smoke_wave + smoke10, or_ts)
# if we used smoke >15 we'd add 164 extra days 
# xtabs(~smoke_wave + smoke15, or_ts)

# small multiples dataset
or_smk <- or_ts %>% 
  select(county, date, wrf_smk_pm_county, krig_smk_pm_county, geo_smk_pm_county) %>% 
  gather(key = smk_method, value = pm, -date, -county)

@@

# data wrangle ----
#first step is to summarise the counts of smoke days in each county 
or_smk_count <- or_ts %>% 
  group_by(county) %>% 
  # summing up smoke days
  summarise(smk_wave_n = sum(smoke_wave), smk_day0 = sum(smoke0), 
            smk_day5 = sum(smoke5), smk_day10 = sum(smoke10),
            smk_day15 = sum(smoke15)) %>% 
  # county needs to be lowercase to join to maps dataframe
  mutate(county = tolower(county))

@@

# outcome name
outcome_name <- "Beta - 2 agonist"

@@

The overall plot shows that when smoke concentration increases 10ug/m^3, the risk to get beta-2 agonist fill will increase about 7.5% significantly.

@@

method_list <- c('Geo-Weighted Smoke')

@@

# outcome name
outcome <- "Beta - 2 agonists"
outcome_name <- "Beta-2 agonists"

@@

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases

@@

# empty matrix (12 x 10 matrix)
point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)

@@

# dataframe for analysis creation
# bind columns back together 
df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
  # remove missing pm values
  filter(!is.na(geo_smk_pm_zip)) %>% 
  # the following code makes sure that the counterfactual values retained are 
  # symetric in that number of obs before = number of obs after
  mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

# second loop to run a model for each pm estimation method
j <- 38

@@

  # create else statement that fills matrix with missing so I still have the row
  # in the final dataframe
  } else {point_estimates[row_n, 2] <- method_list
          point_estimates[row_n, 3] <- 0
          point_estimates[row_n, 4] <- 0
          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement

@@

wrf_geo_age <- combined_point_est_df %>% 
  # subset columns I want to put in to the table
  select(2, 4:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
         # row group by outcome
         rgroup = "Beta - 2 agonist",
         n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Est Method", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(3, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
         ) # end table

@@

# outcome name
outcome <- "Beta - 2 agonists"
outcome_name <- "Beta - 2 agonists"

@@

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases

# new loop for age categories
for(k in 0:2){
  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)

@@

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(geo_smk_pm_zip)) %>% 
    # limit to specific age category
    filter(age_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
    # dataframe is already for the entire fire season, so I don't need to subset anymore

 j <- 38

@@

    # create else statement that fills matrix with missing so I still have the row
    # in the final dataframe
    } else {point_estimates[row_n, 3] <- method_list[row_n]
            point_estimates[row_n, 4] <- 0
            point_estimates[row_n, 5] <- 0
            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
     # end methods loop

@@

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
         # row group by outcome
         rgroup = "Beta - 2 agonists",
         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Est Method", "Age Group", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(4, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
         ) # end table

@@

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_age_fig_gwr.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

@@

The association with smoke PM~2.5~ and health outcomes stratified by sex is shown below.

@@

These results show that there might be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.

```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
method_list <- c('Geo-Weighted Smoke')

@@

  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)

@@

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(geo_smk_pm_zip)) %>% 
    # limit to specific sex category
    filter(sex_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
    # dataframe is already for the entire fire season, so I don't need to subset anymore

  j <- 38
    # variable to model 
    var_name <- colnames(df_analysis[j])

@@

  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
           # row group by outcome
           rgroup = "Beta - 2 agonists",
           n.rgroup = c(rep(2, 1)), # 2 rows for each sex strata for each outcome
           # column headers
           header = c("Est Method", "Sex Strata", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc", # column alignment,
           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
            ) # end table

@@

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_sex_fig_gwr.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

@@

Although the casecrossover study shows a significant association of getting beta-2 agonist pharmacy fills and smoke concentration, we still wonder the association of the pharmacy fills with some lag days. For medication analysis, people could get the pharmacy fills one or some days later than the day they find they have the disease. Then we make 0 (the original day) to 7 lags (one week) to see association of pharmacy fills with different lagged days.

@@

# fill outcome name in first column
fit_mat[,1] <- "Beta - 2 agonists"
# fill in exposure (geo smoke)
fit_mat[,2] <- "geo_smk10"
# fit type
fit_mat[,3] <- "ns"

@@

kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")

@@

# kable
knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
  "degree of freedom best fit by AIC"))
```

@@

When choosing the df of 5, the plot shows that the same day of fill with smoke exposure has the highest significant association. But other lags shows that pharmacy fills with lagged days do not has the significant association with smoke. And the association tends to decrease as lag goes.

```{r DLM for casecrossover df 5, echo = FALSE, message=FALSE, warning=FALSE}
inhaler_geo_lag <- inhaler_ndc_lag_df %>%
  select(159, 160:166) # GWR-method

@@

# the nrow B should equal ncol(pm)
# dim(B)

@@

# coef(fit)
# summary(fit)
# AIC(fit)

@@

# DLestimate

@@

cumulative <- sum(DLestimate$estimate)
# cumulative

cumulative_se <- sqrt(sum(DLvar))
# cumulative_se

cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
# cumulative_CI

cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")

knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))

@@

Then we make the reversed lag to check with the same df of 5. The plot shows the same result as previous analysis.

```{r DLM for casecrossover lag -7 to 0 df 5, echo = FALSE, message=FALSE, warning=FALSE}
inhaler_geo_lag <- inhaler_ndc_lag_df %>%
  select(166:160, 159) # GWR-method

inhaler_temp_lag <- inhaler_ndc_lag_df %>%
  select(156:150, 93) # GWR-method

B <- ns(-7:0, df =5, intercept=TRUE)

# the nrow B should equal ncol(pm)
# dim(B)

# coef(fit)
# summary(fit)
# AIC(fit)

# DLestimate

DLestimate$lag <- -7:0
p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
p <- p + geom_line(aes(x=lag), linetype=2, size=2)
p <- p + theme_bw()
p

cumulative <- sum(DLestimate$estimate)
# cumulative

cumulative_se <- sqrt(sum(DLvar))
# cumulative_se

cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
# cumulative_CI

cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")

knitr::kable(cum, caption = paste0("Distributed lag -7 to 0 spline estimates, se and CI"))

@@

```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE, eval=FALSE}
read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
or_ndc_claim <- read_csv(read_path7)

@@

```{r time stratified zip for casecrossover lag -7 to 0 df, echo=FALSE, include=FALSE, eval=FALSE}
read_path10 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
inhaler_ndc_lag15 <- read_csv(read_path10)
# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis

@@

@@

# outcome name
outcome <- "Beta - 2 agonists"

@@

# extract pm values and divide by 10 and ordered
# which(colnames(inhaler_ndc)=="geo_smk_pm_lag1_zip") # 136
# which(colnames(inhaler_ndc)=="wrf_temp_lag1_zip") # 150
pm_estimates_df <- inhaler_ndc[, 136:142]/10  # create 10 unit increases
temp_estimates_df <- inhaler_ndc[, 150:156]  # temperature

@@

tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
         # row group by outcome
         rgroup = "Beta - 2 agonists",
         n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Lag", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(3, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
         ) # end table

@@

rr_df[i,3:5] <- rr_mp
}

kable(rr_df, caption = "Unconstrained lag for time series")

@@

(not shown)

@@

### 4. Distributed lag -7 to 0 for time series

(not shown)

```{r DLM for casecrossover lag -7 to 0 temp, echo = FALSE, message=FALSE, warning=FALSE, eval = FALSE}
